---
title: "Ted Talk"
output: html_notebook
---

Some of the questions that can be answered with this dataset: 1. How is each TED Talk related to every other TED Talk? 2. Which are the most viewed and most favorited Talks of all time? Are they mostly the same? What does this tell us? 3. What kind of topics attract the maximum discussion and debate (in the form of comments)? 4. Which months are most popular among TED and TEDx chapters? 5. Which themes are most popular amongst TEDsters?

Read the data files

```{r}
library(readr)
ted <- read_csv("ted-talks/ted_main.csv")
head(ted)
```

```{r}
transcript <- read_csv("ted-talks/transcripts.csv")
head(transcript)
```

See the information of both datasets

```{r}
summary(ted)
```

```{r}
summary(transcript)
```

Check the dimensions of the data
```{r}
dim(ted)
dim(transcript)
```

Features Available
- name: The official name of the TED Talk. Includes the title and the speaker.
- title: The title of the talk
- description: A blurb of what the talk is about.
- main_speaker: The first named speaker of the talk.
- speaker_occupation: The occupation of the main speaker.
- num_speaker: The number of speakers in the talk.
- duration: The duration of the talk in seconds.
- event: The TED/TEDx event where the talk took place.
- film_date: The Unix timestamp of the filming.
- published_date: The Unix timestamp for the publication of the talk on TED.com
- comments: The number of first level comments made on the talk.
- tags: The themes associated with the talk.
- languages: The number of languages in which the talk is available.
- ratings: A stringified dictionary of the various ratings given to the talk (inspiring, fascinating, jaw dropping, etc.)
- related_talks: A list of dictionaries of recommended talks to watch next.
- url: The URL of the talk.
- views: The number of views on the talk.
I'm just going to reorder the columns in the order I've listed the features for my convenience (and OCD).

Let's change the UNIX timestamps to human readable date

```{r}
library(anytime)
ted$film_date <- anydate(ted$film_date)
ted$published_date <- anydate(ted$published_date)
```

Exploratory Data Analysis
- Which talks are most and least viewed?



Which talks are most and least viewed?

```{r}
library(dplyr)

ten_talks <- arrange(ted, views)
keeps <- c("title", "main_speaker", "views", "speaker_occupation", "film_date")
ten_talks <- subset(ten_talks, select = c("title", "main_speaker", "views", "speaker_occupation", "film_date"))

last_10_talks <- ten_talks[0:10,]
last_10_talks

ten_talks <- arrange(ted, desc(views))
ten_talks <- subset(ten_talks, select = c("title", "main_speaker", "views", "speaker_occupation", "film_date"))
top_10_talks <- ten_talks[0:10,]
top_10_talks
```

Observations
Ken Robinson's talk on Do Schools Kill Creativity? is the most popular TED Talk of all time with 47.2 million views.
Also coincidentally, it is also one of the first talks to ever be uploaded on the TED Site (the main dataset is sorted by published date).
Robinson's talk is closely followed by Amy Cuddy's talk on Your Body Language May Shape Who You Are.
There are only 2 talks that have surpassed the 40 million mark and 4 talks that have crossed the 30 million mark.
Let us make a bar chart to visualise these 15 talks in terms of the number of views they garnered.

```{r fig.width=7, fig.height=4, echo=FALSE}
library(ggplot2)
library(Rmisc)
tilt_theme <- theme(axis.text.x=element_text(angle=45, hjust=1))
p1 <- ggplot(data = last_10_talks, aes(main_speaker, views)) +
      geom_bar(stat = "identity") +
      geom_text(aes(label=views), vjust=1.6, color="white", size=3) +
      tilt_theme
p2 <- ggplot(data = top_10_talks, aes(main_speaker, views)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = views), vjust = 1.6, color = "white", size = 3) +
      tilt_theme
multiplot(p1, p2, cols=2)
```

Distribution of Views

```{r}
ted_small <- subset(ted, views < 0.4e7)
p3 <- ggplot(ted, aes(x=views)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666")
p4 <- ggplot(ted_small, aes(x=views)) +
  geom_histogram(aes(y=..density..), colour="black", fill="white")+
  geom_density(alpha=.2, fill="#FF6666")
multiplot(p3, p4, cols=2)
```

```{r}
library(Hmisc)
describe(ted$views)
```

The average number of views on TED Talks in 1.6 million. and the median number of views is 1.12 million. This suggests a very high average level of popularity of TED Talks. We also notice that the majority of talks have views less than 4 million. 

Comments
Let's analyze comments on the talks
```{r}
describe(ted$comments)
sd(ted$comments)
```

Observations
- On average, there are 191.5 comments on every TED Talk. Assuming the comments are constructive criticism, we can conclude that the TED Online Community is highly involved in discussions revolving TED Talks.
- There is a huge standard deviation associated with the comments. In fact, it is even larger than the mean suggesting that the measures may be sensitive to outliers. We shall plot this to check the nature of the distribution.
- The minimum number of comments on a talk is 2 and the maximum is 6404. The range is 6402.. The minimum number, though, may be as a result of the talk being posted extremely recently.

```{r}
ted_small <- subset(ted, comments < 500)
p5 <- ggplot(ted, aes(x=comments)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666")
p6 <- ggplot(ted_small, aes(x=comments)) +
  geom_histogram(aes(y=..density..), colour="black", fill="white")+
  geom_density(alpha=.2, fill="#FF6666")
multiplot(p5, p6, cols=2)
```

From the plot above, we can see that the bulk of the talks have fewer than 500 comments. This clearly suggests that the mean obtained above has been heavily influenced by outliers. This is possible because the number of samples is only 2550 talks.

Another question that I am interested in is if the number of views is correlated with the number of comments. We should think that this is the case as more popular videos tend to have more comments. Let us find out.

```{r}
scatter1 <- qplot(views,comments, data=ted)  + 
         scale_x_continuous(limits=c(min(ted$views),max(ted$views))) + 
         scale_y_continuous(limits=c(min(ted$comments),max(ted$comments))) + 
         geom_rug(col=rgb(.5,0,0,alpha=.2))
scatter1
```

```{r}
vc_cor <- cor(ted$views, ted$comments)
print("Correlation between Views and Comments")
vc_cor
```

As the scatterplot and the correlation matrix show, the pearson coefficient is slightly more than 0.5. This suggests a medium to strong correlation between the two quantities. This result was pretty expected as mentioned above. Let us now check the number of views and comments on the 10 most commented TED Talks of all time.

```{r}
ten_talks <- arrange(ted, comments)
keeps <- c("title", "main_speaker", "views", "comments")
ten_talks <- subset(ten_talks, select = keeps)

last_10_talks <- ten_talks[0:10,]
last_10_talks

ten_talks <- arrange(ted, desc(comments))
ten_talks <- subset(ten_talks, select = keeps)
top_10_talks <- ten_talks[0:10,]
top_10_talks
```

```{r fig.width=7, fig.height=4, echo=FALSE}

tilt_theme <- theme(axis.text.x=element_text(angle=45, hjust=1))
p7 <- ggplot(data = last_10_talks, aes(main_speaker, comments)) +
      geom_bar(stat = "identity") +
      geom_text(aes(label=comments), vjust=1.6, color="white", size=3) +
      tilt_theme
p8 <- ggplot(data = top_10_talks, aes(main_speaker, comments)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = comments), vjust = 1.6, color = "white", size = 3) +
      tilt_theme
multiplot(p7, p8, cols=2)

```

As can be seen above, Richard Dawkins' talk on Militant Atheism' generated the greatest amount of discussion and opinions despite having significantly lesser views than Ken Robinson's talk, which is second in the list. This raises some interesting questions.

Which talks tend to attract the largest amount of discussion?

To answer this question, we will define a new feature discussion quotient which is simply the ratio of the number of comments to the number of views. We will then check which talks have the largest discussion quotient.

```{r}
ted$dis_quo <- ted$comments/ted$views
ten_talks <- arrange(ted, dis_quo)
keeps <- c("title", "main_speaker", "views", "comments", "dis_quo")
ten_talks <- subset(ten_talks, select = keeps)

last_10_talks <- ten_talks[0:10,]
last_10_talks

ten_talks <- arrange(ted, desc(dis_quo))
ten_talks <- subset(ten_talks, select = keeps)
top_10_talks <- ten_talks[0:10,]
top_10_talks
```

This analysis has actually raised extremely interesting insights. Half of the talks in the top 10 are on the lines of Faith and Religion. I suspect science and religion is still a very hotly debated topic even in the 21st century. We shall come back to this hypothesis in a later section.

The most discusses talk, though, is The Case for Same Sex Marriage (which has religious undertones). This is not that surprising considering the amount of debate the topic caused back in 2009 (the time the talk was filmed).

########################

Analyzing TED talks by the month and the year

TED (especially TEDx) Talks tend to occur all throughout the year. Is there a hot month as far as TED is concerned? In other words, how are the talks distributed throughout the months since its inception? Let us find out.

```{r}
ted$filming_month <- as.numeric(format(ted$film_date, format= "%m"))
ted$filming_year <- as.numeric(format(ted$film_date, format = "%Y"))
month_df <- as.data.frame(table(ted$filming_month))
colnames(month_df) <- c("Month", "Talks")
month_df$Month <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
#Turn your 'Month' column into a character vector
month_df$Month <- as.character(month_df$Month)
#Then turn it back into an ordered factor
month_df$Month <- factor(month_df$Month, levels=unique(month_df$Month))
p8 <- ggplot(data = month_df, aes(Month, Talks)) +
      geom_bar(position = "dodge", stat = "identity", color = "blue", fill="blue") +
      geom_text(aes(label = Talks), vjust = 1.6, color = "white", size = 3) +
      tilt_theme
p8

```

February is clearly the most popular month for TED Conferences whereas August and January are the least popular. February's popularity is largely due to the fact that the official TED Conferences are held in February. Let us check the distribution for TEDx talks only.

```{r}
tedx <- filter(ted, grepl("TEDx", ted$event, fixed = TRUE))
month_df <- as.data.frame(table(tedx$filming_month))
colnames(month_df) <- c("Month", "Talks")
month_df$Month <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
#Turn your 'Month' column into a character vector
month_df$Month <- as.character(month_df$Month)
#Then turn it back into an ordered factor
month_df$Month <- factor(month_df$Month, levels=unique(month_df$Month))
p9 <- ggplot(data = month_df, aes(Month, Talks)) +
      geom_bar(position = "dodge", stat = "identity", color = "red", fill="red") +
      geom_text(aes(label = Talks), vjust = 1.6, color = "white", size = 3) +
      tilt_theme
p9
```

As far as TEDx talks are concerned, November is the most popular month. However, we cannot take this result at face value as very few of the TEDx talks are actually uploaded to the TED website and therefore, it is entirely possible that the sample in our dataset is not at all representative of all TEDx talks. A slightly more accurate statement would be that the most popular TEDx talks take place the most in October and November.

The next question I'm interested in is the most popular days for conducting TED and TEDx conferences. The tools applied are very sensible to the procedure applied for months.

```{r}
ted$film_weekday <- weekdays(as.Date(ted$film_date))
weekday_df <- as.data.frame(table(ted$film_weekday))

weekday_df

colnames(weekday_df) <- c("Weekday", "Talks")
weekday_df <- weekday_df[c(4,2,6,7,5,1,3),]
#Turn your 'Month' column into a character vector
weekday_df$Weekday <- as.character(weekday_df$Weekday)
#Then turn it back into an ordered factor
weekday_df$Weekday <- factor(weekday_df$Weekday, levels=unique(weekday_df$Weekday))
p10 <- ggplot(data = weekday_df, aes(factor(Weekday), Talks)) +
      geom_bar(position = "dodge", stat = "identity", color = "green", fill="green") +
      geom_text(aes(label = Talks), vjust = 1.6, color = "black", size = 3) +
      tilt_theme
p10
```

The distribution of days is almost a bell curve with Wednesday and Thursday being the most popular days and Sunday being the least popular. This is pretty interesting because I was of the opinion that most TED Conferences would happen sometime in the weekend.

Let us now visualize the number of TED talks through the years and check if our hunch that they have grown significantly is indeed true.

```{r}
year_df <- as.data.frame(table(ted$filming_year))
year_df
colnames(year_df) <- c("Year", "Talks")
#Turn your 'Month' column into a character vector
#month_df$Month <- as.character(month_df$Month)
#Then turn it back into an ordered factor
#month_df$Month <- factor(month_df$Month, levels=unique(month_df$Month))
p11 <- ggplot(data = year_df, aes(Year, Talks, group = 1)) +
      geom_line(color = "blue") +
      geom_point() +
      geom_text(aes(label = Talks), vjust = 1.6, color = "black", size = 3) +
      tilt_theme
p11
```

Obervations
- As expected, the number of TED Talks have gradually increased over the years since its inception in 1984.
- There was a sharp increase in the number if talks in 2009. It might be interesting to know the reasons behind 2009 being the tipping point where the number of talks increased more than twofold.
- The number of talks have been pretty much constant since 2009.

Finally, to put it all together, let us construct a heat map that shows us the number of talks by month and year. This will give us a good summary of the distribution of talks.

```{r}

keeps <- c("filming_month", "filming_year")
heatmap_df <- ted %>% subset(select = keeps) %>% arrange(filming_year)
heatmap_df <- as.data.frame(table(heatmap_df))
heatmap_df$filming_month <- month.abb[heatmap_df$filming_month]
heatmap_df$filming_month <- as.character(heatmap_df$filming_month)

heatmap_df$filming_month <- factor(heatmap_df$filming_month, levels=unique(month_df$Month))
#heatmap_df 

ggplot(heatmap_df, aes(filming_year, factor(filming_month))) +
  geom_tile(aes(fill = Freq), color = "white") +
  geom_text(aes(fill = heatmap_df$Freq, label = heatmap_df$Freq)) +
  scale_fill_gradient2(low = "darkred", 
                       mid = "white", 
                       high = "midnightblue", 
                       midpoint = round(mean(heatmap_df$Freq))) +
  ylab("Year") +
  xlab("Months") +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 12),
        plot.title = element_text(size=16),
        axis.title=element_text(size=14,face="bold"),
        axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(fill = "Freq")

```

######
Ted Speakers

In this section, we will try and gain insight about all the amazing speakers who have managed to inspire millions of people through their talks on the TED Platform. The first question we shall ask in this section is who are the most popular TED Speakers. That is, which speakers have given the most number of TED Talks.

```{r}
speaker_df <- as.data.frame(table(ted$main_speaker))
speaker_df <- arrange(speaker_df, desc(Freq))
speaker_df
```

Hans Rosling, the Swiss Health Professor is clearly the most popular TED Speaker, with more than 9 appearances on the TED Forum. Juan Enriquez comes a close second with 7 appearances. Rives and Marco Tempest have graced the TED platform 6 times.

Which occupation should you choose if you want to become a TED Speaker? Let us have a look what kind of people TED is most interested in inviting to its events.

```{r}
occupation_df <- as.data.frame(table(ted$speaker_occupation))
occupation_df <- arrange(occupation_df, desc(Freq))
occupation_df <- head(occupation_df, 10)
colnames(occupation_df) <- c("Occupation", "Appearances")
occupation_df$Occupation <- as.character(occupation_df$Occupation)
occupation_df$Occupation <- factor(occupation_df$Occupation, levels=occupation_df$Occupation)
p12 <- ggplot(data = occupation_df, aes(factor(Occupation), Appearances)) +
      geom_bar(position = "dodge", stat = "identity", color = "cyan", fill="cyan") +
      geom_text(aes(label = Appearances), vjust = 1.6, color = "black", size = 3) +
      tilt_theme
p12
```

Observations
- Writers are the most popular with more than 45 speakers identifying themselves as the aforementioned.
- Artists and Designers come a distant second with around 35 speakers in each category.
- This result must be taken with a pinch of salt as a considerable number of speakers identify themselves with multiple professions (for example, writer/entrepreneur). 

Do some professions tend to attract a larger number of viewers? Do answer this question let us visualise the relationship between the top 10 most popular professions and the views that garnered in the form of a box plot.


```{r}
top_occupation_views <- filter(ted, ted$speaker_occupation %in% occupation_df$Occupation)
p13 <- ggplot(top_occupation_views, aes(x=speaker_occupation, y=views)) + 
  geom_boxplot() +
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  tilt_theme
p13 
```

On average, out of the top 10 most popular professions, Psychologists tend to garner the most views. Writers have the greatest range of views between the first and the third quartile..

Finally, let us check the number of talks which have had more than one speaker.

```{r}
speaker_count <- table(ted$num_speaker)
speaker_count
```

Almost every talk has just one speaker. There are close to 50 talks where two people shared the stage. The maximum number of speakers to share a single stage was 5. I suspect this was a dance performance. Let's have a look.

```{r}
filter(ted, ted$num_speaker==5)
```

My hunch was correct. It is a talk titled A dance to honor Mother Earth by Jon Boogz and Lil Buck at the TED 2017 Conference.

#################
TED Events
Which TED Events tend to hold the most number of TED.com upload worthy events? We will try to answer that question in this section.


```{r}
events_df <- ted %>% select(c("event")) %>% group_by("event")
events_df <- as.data.frame(table(events_df))
events_df <- arrange(events_df, desc(Freq))
events_df <- head(events_df, 10)

p14 <- ggplot(data = events_df, aes(event, Freq)) +
      geom_bar(position = "dodge", stat = "identity", color = "blue", fill="blue") +
      geom_text(aes(label = Freq), vjust = 1.6, color = "white", size = 3) +
      tilt_theme
p14
```

As expected, the official TED events held the major share of TED Talks published on the TED.com platform. TED2014 had the most number of talks followed by TED2009. There isn't too much insight to be gained from this.

##############
TED Languages
One remarkable aspect of TED Talks is the sheer number of languages in which it is accessible. Let us perform some very basic data visualisation and descriptive statistics about languages at TED.

```{r}
describe(ted$languages)
```

On average, a TED Talk is available in 27 different languages. The maximum number of languages a TED Talk is available in is a staggering 72. Let us check which talk this is.

```{r}
filter(ted, ted$languages == 72)
```

The most translated TED Talk of all time is Matt Cutts' Try Something New in 30 Days. The talk does have a very universal theme of exploration. The sheer number of languages it's available in demands a little more inspection though as it has just over 8 million views, far fewer than the most popular TED Talks.

Finally, let us check if there is a correlation between the number of views and the number of languages a talk is availbale in. We would think that this should be the case since the talk is more accessible to a larger number of people but as Matt Cutts' talk shows, it may not really be the case.

```{r}
scatter2 <- qplot(languages, views, data=ted)  + 
         scale_y_continuous(limits=c(min(ted$views),max(ted$views))) + 
         scale_x_continuous(limits=c(min(ted$languages),max(ted$languages))) + 
         geom_rug(col=rgb(.5,0,0,alpha=.2))
scatter2
```

```{r}
cor(ted$languages, ted$views)
```

The Pearson coefficient is 0.38 suggesting a medium correlation between the aforementioned quantities.

##############
TED Themes
In this section, we will try to find out the most popular themes in the TED conferences. Although TED started out as a conference about technology, entertainment and design, it has since diversified into virtually every field of study and walk of life. It will be interesting to see if this conference with Silicon Valley origins has a bias towards certain topics.

To answer this question, we need to wrangle our data in a way that it is suitable for analysis. More specifically, we need to split the related_tags list into separate rows.

```{r}
perfect_tag <- function(x){
  x <- unlist(strsplit(x, "'"))
  val = x[2]
  for (i in 3:length(x))
    if (nchar(x[i]) >2)
      val <- c(val, x[i])
    return (val)
  
}
ted$processed_tags <-  lapply(ted$tags, perfect_tag)
```

```{r}
processed_tags <- ted$processed_tags
length(processed_tags)
processed_tags <- unlist(processed_tags, recursive=FALSE)
length(processed_tags)
processed_tags <- as.data.frame(table(processed_tags))
processed_tags <- arrange(processed_tags, desc(Freq))
head(processed_tags, 10)
```

```{r}
processed_tags$processed_tags <- as.character(processed_tags$processed_tags)
processed_tags$processed_tags <- factor(processed_tags$processed_tags, levels=processed_tags$processed_tags)
p15 <- ggplot(data = head(processed_tags, 10), aes(processed_tags, Freq)) +
      geom_bar(position = "dodge", stat = "identity", color = "blue", fill="blue") +
      geom_text(aes(label = Freq), vjust = 1.6, color = "white", size = 3) +
      tilt_theme
p15
```

As may have been expected, Technology is the most popular topic for talks. The other two original factions, Design and Entertainment, also make it to the list of top 10 themes. Science and Global Issues are the second and the third most popular themes respectively.

The next question I want to answer is the trends in the share of topics of TED Talks across the world. Has the demand for Technology talks increased? Do certain years have a disproportionate share of talks related to global issues? Let's find out!

We will only be considering the top 7 themes, excluding TEDx and talks after 2009, the year when the number of TED Talks really peaked.

```{r}
factor_fn <- function(x){
  return(factor(x, levels = x))
}
ted$processed_tags <- lapply(ted$processed_tags, unlist, recursive = FALSE)
ted$processed_tags <- lapply(ted$processed_tags, factor_fn)
rows_to_keep <- rep(TRUE, length(ted$processed_tags))
for (i in 1:length(ted$processed_tags)){
  if (any(unlist(ted$processed_tags[i]) %in% processed_tags$processed_tags[1:10])){
    rows_to_keep[i] <- TRUE} else {
    rows_to_keep[i] <- FALSE}
}
#pop_theme_talks <- filter(pop_theme_talks, any(ted$processed_tags %in% processed_tags$processed_tags[1:10]))
#pop_theme_talks
```


```{r}
pop_theme_talks <- ted[rows_to_keep,]
pop_theme_talks <- filter(pop_theme_talks, filming_year>2008)
```

51 input



